\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2022-02-03 4-4-netcdf.Rmd
netcdf4 section done

Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`. It maps each combination of latitude and longitude into a unit.

@@

Since each combination of latitude and longitude grid is regarded as a site, the number of site quickly goes up when mapping on a global scale.

  - For a southern hemisphere bounding box of [-5, -180, -90, 180], 491,040 id, 8 timestamp, 15MB -> 11 sec
  - For an Australia bounding box of [-12, 115, -43, 153], 19125 id, 1,827 times tamp, 133 MB -> 1 mins .

This shows a computation on the time dimension is cheaper than the spatial dimension. Argument `long_range` and `lat_range` are available to subset to a looser grid when reading in global level resolution. `era5-pressure.nc` has 0.25 degree resolution and to read in the data in whole degree, use:

```{r eval = FALSE, echo = TRUE}
dt <- as_cubble(raw, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure 19 in @hersbach2020era5 from the ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format. It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr]. Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Readers who are interested in the details of  can refer to @simmons2020global and @simmons2005ecmwf.

```{r netcdf, out.width="100%", fig.cap = "sdf"}
knitr::include_graphics(here::here("figures/netcdf.png"))
```

>>>>

Cubble provides an `as_cubble()` method to coerce the `ncdf4` class from the `ncdf4` package into a `cubble`. It maps each combination of longitude and latitude into an `id` (the `key`) and nests time and variables in the nested form:

@@

Memory limit with Netcdf data in cubble depends on longitude grid point x latitude grid point x time grid point x number of variable. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for 3 variables in one year and spatial grid can be reduced to trade for longer time period and more variables. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or global bounding box [-180, -90, 180, -90] at 1 degree resolution. Subsetting longitude and latitude grid is available through `long_range` and `lat_range` if the Netcdf file has finer resolution than needed.

```{r eval = FALSE, echo = TRUE}
# Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
# at 0.25 degree resolution and subset it to have 
# 1 degree resolution:
dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
                long_range = seq(-180, 180, 1), 
                lat_range = seq(-90, 90, 1))
```

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in @hersbach2020era5 using ERA5 data. ERA5 data [@hersbach2020era5] is the latest reanalysis of global atmosphere, land surface, and ocean waves from 1950 onwards and is available in the NetCDF format. It can be directly downloaded from the [Copernicus Climate Data Store (CDS)](https://cds.climate.copernicus.eu/cdsapp#!/dataset/reanalysis-era5-pressure-levels?tab=overview) website or programmatically via R package `ecmwfr`[@ecwmfr]. Variable Specific humidity and geopotential are queried on the 10 hPa pressure level for four dates: 2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded, the data can be read into a cubble using the code above. Readers who are interested in the details of can refer to @simmons2020global and @simmons2005ecmwf.

```{r netcdf, out.width="100%", fig.cap = "A reproduction of the second row of Figure 19 in  Hersbach (2020)."}
knitr::include_graphics(here::here("figures/netcdf.png"))
```

### 2022-02-03 main.tex
netcdf4 section done

Cubble provides an \texttt{as\_cubble()} method to coerce the
\texttt{ncdf4} class from the \texttt{ncdf4} package into a
\texttt{cubble}. It maps each combination of latitude and longitude into
a unit.

@@

Since each combination of latitude and longitude grid is regarded as a
site, the number of site quickly goes up when mapping on a global scale.

\begin{itemize}
\tightlist
\item
  For a southern hemisphere bounding box of {[}-5, -180, -90, 180{]},
  491,040 id, 8 timestamp, 15MB -\textgreater{} 11 sec
\item
  For an Australia bounding box of {[}-12, 115, -43, 153{]}, 19125 id,
  1,827 times tamp, 133 MB -\textgreater{} 1 mins .
\end{itemize}

This shows a computation on the time dimension is cheaper than the
spatial dimension. Argument \texttt{long\_range} and \texttt{lat\_range}
are available to subset to a looser grid when reading in global level
resolution. \texttt{era5-pressure.nc} has 0.25 degree resolution and to
read in the data in whole degree, use:

\begin{CodeChunk}
\begin{CodeInput}
R> dt <- as_cubble(raw, vars = c("q", "z"), 
+                 long_range = seq(-180, 180, 1), 
+                 lat_range = seq(-90, 90, 1))
\end{CodeInput}
\end{CodeChunk}

Here an example in Figure \ref{fig:netcdf} is given to reproduce Figure
19 in \citet{hersbach2020era5} from the ERA5 data. ERA5 data
\citep{hersbach2020era5} is the latest reanalysis of global atmosphere,
land surface, and ocean waves from 1950 onwards and is available in the
NetCDF format. It can be directly downloaded from the
\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
Climate Data Store (CDS)} website or programmatically via R package
\texttt{ecmwfr}\citep{ecwmfr}. Variable Specific humidity and
geopotential are queried on the 10 hPa pressure level for four dates:
2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04. Readers who are
interested in the details of can refer to \citet{simmons2020global} and
\citet{simmons2005ecmwf}.

@@

\caption[sdf]{sdf}\label{fig:netcdf}
\end{figure}
\end{CodeChunk}

>>>>

Cubble provides an \texttt{as\_cubble()} method to coerce the
\texttt{ncdf4} class from the \texttt{ncdf4} package into a
\texttt{cubble}. It maps each combination of longitude and latitude into
an \texttt{id} (the \texttt{key}) and nests time and variables in the
nested form:

@@

Memory limit with Netcdf data in cubble depends on longitude grid point
x latitude grid point x time grid point x number of variable. Cubble can
handle slightly more than 300 x 300 (longitude x longitude) grid points
for 3 variables in one year and spatial grid can be reduced to trade for
longer time period and more variables. A 300 by 300 spatial grid can be
a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or
global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution.
Subsetting longitude and latitude grid is available through
\texttt{long\_range} and \texttt{lat\_range} if the Netcdf file has
finer resolution than needed.

\begin{CodeChunk}
\begin{CodeInput}
R> # Assume `my_ncdf` has bounding box of [-180, -90, 180, -90] 
R> # at 0.25 degree resolution and subset it to have 
R> # 1 degree resolution:
R> dt <- as_cubble(my_ncdf, vars = c("q", "z"), 
+                 long_range = seq(-180, 180, 1), 
+                 lat_range = seq(-90, 90, 1))
\end{CodeInput}
\end{CodeChunk}

Figure \ref{fig:netcdf} gives an example of reproducing Figure 19 in
\citet{hersbach2020era5} using ERA5 data. ERA5 data
\citep{hersbach2020era5} is the latest reanalysis of global atmosphere,
land surface, and ocean waves from 1950 onwards and is available in the
NetCDF format. It can be directly downloaded from the
\href{https://cds.climate.copernicus.eu/cdsapp\#!/dataset/reanalysis-era5-pressure-levels?tab=overview}{Copernicus
Climate Data Store (CDS)} website or programmatically via R package
\texttt{ecmwfr}\citep{ecwmfr}. Variable Specific humidity and
geopotential are queried on the 10 hPa pressure level for four dates:
2002-09-22, 2002-09-26, 2002-09-30, and 2002-10-04 adn once downloaded,
the data can be read into a cubble using the code above. Readers who are
interested in the details of can refer to \citet{simmons2020global} and
\citet{simmons2005ecmwf}.

@@

\caption[A reproduction of the second row of Figure 19 in  Hersbach (2020)]{A reproduction of the second row of Figure 19 in  Hersbach (2020).}\label{fig:netcdf}
\end{figure}
\end{CodeChunk}
